### 2019-09-09 ic-journal/response-letter.tex

  It should be ``via''.

\item \begin{quote}
    L184: closes -> completes
  \end{quote}
  Done.

\item \begin{quote}
    L194: ‘the definition should not be recursive’?
  \end{quote}

  We meant the definition should not be \emph{coinductive}, just like the
  case of $\rapprox$. Notice that, this is just an informal idea when searching for
  a possible definiton of rooted contraction, and the definition we
  found at last is indeed not coinductive (nor recursive).

\item \begin{quote}
    Definition 3.11: Similar to the remark about Definition 2.2.
  \end{quote}
  You meant removing also the comma (period) here? But it's different
  here: the definition ends at clause (2), an end period is added with
  an ``enspace'' distance.

### 2019-09-09 ic-journal/contraction.tex
Finished revising the first past

The constraints on the unique-solution Theorem~\ref{t:Mil89} can be 
weakened if we move from equations to a special kind of inequations called
  \emph{contractions}.

Intuitively, the bisimilarity contraction $\mcontrBIS$ is a preorder in which 
$P \mcontrBIS Q  $  holds  if $P \wb Q$ and, in addition, 
$Q$ has the \emph{possibility} of being at least as efficient as $P$ (as far as
$\tau$-actions performed). 
Process $Q$, however, may be nondeterministic and may have other ways
of doing the same work, and these could be  slow (i.e., involving
more $\tau$-steps than those performed by $P$).
% Thus, in contrast with expansion,  we cannot really say that `$Q$ is more efficient than
% $P$'.

\begin{definition}%[bisimulation contraction]
\label{d:BisCon}
A process relation ${\R}$ 
 is a \emph{(bisimulation) contraction} if, whenever
 $P\RR Q$, %for all $\mu$

@@

In the first clause $Q$ is required to match $P$'s challenge
transition with at most one transition.
This makes sure that $Q$ is capable of mimicking $P$'s
work at least as efficiently as $P$. 
In contrast, the second clause of Definition~\ref{d:BisCon}, on the
challenges from $Q$, entirely ignores efficiency: it is the same
clause of  weak bisimulation~---the final derivatives are \hl{indeed} required
to be related  by $\wb$, rather than by $\R$.

Bisimilarity  contraction is coarser than 
the \emph{expansion relation} $\expa$~\cite{arun1992efficiency,sangiorgi2015equations}.
This is a preorder \hl{that is} widely used in proof techniques for bisimilarity and
intuitively refines bisimilarity by 
formalising the idea of `efficiency' between processes.
Clause (1) is the same in the two
preorders. But in clause (2) expansion uses 
$P \Arr \mu P'$, rather than $P \Arcap \mu P'$; 
 moreover with
contraction the final derivatives are simply required to be bisimilar.
An expansion 
$P \expa Q$
tells us  that $Q$ is always at least as efficient as $P$, whereas  the
 contraction $P \mcontrBIS Q$  just says that $Q$ has the possibility of
being at least as efficient as $P$.

% \iflong\newDS
% Note the use of $\wb$, in place of $\R$, in the second clause:
%  we are only interested to know that $Q$ may  mimic $P$'s
% more in an efficient manner (first clause); we do not care about
% efficiency
% on the
% challenges proposed by $Q$,   which are handled using ordinary
% bisimilarity.
% \fi

\begin{definition}%[expansion]
\label{d:expa}
A process relation ${\R}$
  is an \emph{expansion} if, whenever
we have $P\RR Q$, for all $\mu$
 \begin{enumerate}
 \item   $P \arr\mu P'$ implies that there is $Q'$ with $Q \arcap \mu  Q'$
  and $P' \RR Q'$;
 \item $Q \arr\mu Q'$   implies that there is $P'$ with $P \Arr \mu P'$ and $P' \RR Q'$.
 \end{enumerate}
\hl{Two processes $P$ and $Q$ are in the \emph{bisimilarity
    expansion}}, written as $P \expa Q$ ($P$ \emph{expands} $Q$),
 if $P \RR Q$ for some expansion $\R$.
 \end{definition}

>>>>

Milner's ``unique solution of equations'' theorem for $\wb$
(Theorem~\ref{t:Mil89})
\hl{brings a new proof technique for proving (weak) bisimilarities. However, it has
severe limitations: the equations must be guarded and sequential (plus
all sums where equation variables appear must be guarded sums).
This greatly limits the usefulness of the technique, since
the occurrences of other operators above the variables, such as parallel
composition and restriction, in general cannot be removed.}
The constraints of Theorem~\ref{t:Mil89}, however, can be greatly
weakened if we move from equations to a special kind of inequations called
  \emph{contractions}.

Intuitively, the bisimilarity contraction $\mcontrBIS$ is a preorder
such that, $P \mcontrBIS Q$ holds if $P \wb Q$ and, in addition, 
$Q$ \emph{has the possibility of being at least as efficient as $P$} (as far as
$\tau$-actions are performed).
Process $Q$, however, may be non-deterministic and may have other ways
of doing the same work, and these could be slow (i.e., involving
more $\tau$-\hl{actions} than those performed by $P$).
% Thus, in contrast with expansion,  we cannot really say that `$Q$ is more efficient than
% $P$'.

\begin{definition}[contraction]
\label{d:BisCon}
A process relation ${\R}$ 
 is a \emph{(bisimulation) contraction} if, whenever
 $P\RR Q$, %for all $\mu$

@@

In the first clause $Q$ is required to match $P$'s challenge
transition with at most one transition.
This makes sure that $Q$ is capable of mimicking % verb: mimic
$P$'s work at least as efficiently as $P$. 
In contrast, the second clause of Definition~\ref{d:BisCon}, on the
challenges from $Q$, entirely ignores efficiency: it is the same
clause of  weak bisimulation~---the final derivatives are \hl{indeed} required
to be related  by $\wb$, rather than by $\R$.

Bisimilarity contraction is coarser than 
the \emph{expansion relation}
$\expa$~\cite{arun1992efficiency,sangiorgi2015equations}:
\begin{definition}%[expansion]
\label{d:expa}
A process relation ${\R}$
  is an \emph{expansion} if, whenever
we have $P\RR Q$, for all $\mu$
 \begin{enumerate}
 \item   $P \arr\mu P'$ implies that there is $Q'$ with $Q \arcap \mu  Q'$
  and $P' \RR Q'$;
 \item $Q \arr\mu Q'$   implies that there is $P'$ with $P \Arr \mu P'$ and $P' \RR Q'$.
 \end{enumerate}
\hl{Two processes $P$ and $Q$ are in the \emph{bisimilarity
    expansion}}, written as $P \expa Q$ ($P$ \emph{expands} $Q$),
 if $P \RR Q$ for some expansion $\R$.
 \end{definition}
The expansion preorder is widely used in proof techniques for bisimilarity and
intuitively refines bisimilarity by 
formalising the idea of ``efficiency'' between processes.
Clause (1) is the same in the \hl{both}
preorders, while in clause (2) expansion uses 
$P \Arr \mu P'$, rather than $P \Arcap \mu P'$; 
moreover in clause (2) of Def.~\ref{d:BisCon} the final derivatives
are simply required to be bisimilar ($P' \wb Q'$).
\hl{Intuitively, $P \expa Q$ holds if $P\wb Q$ and, in addition, \emph{$Q$
  is always at least as efficient as $P$}.}

### 2019-09-09 ic-journal/equa.tex
Finished revising the first past

\begin{definition}[equations] % Def 3.1
Assume that, for each $i$ of 
 a countable indexing set $I$, we have variables $X_i$, and expressions
$E_i$ containing variables up to $\til X$. Then 
$\{ X_i = E_i\}_{i\in I}$ is 
  a \emph{system of equations}. (There is one equation $E_i$ for each variable $X_i$.)
\end{definition}

@@

\begin{definition}[solutions and unique solutions]
Suppose $\{ X_i = E_i\}_{i\in I}$ is a system of equations: 
\begin{itemize}
\item
 $\til P$ is a \emph{solution of the system of equations (for $\wb$)} 
if for each $i$ it holds that $P_i \wb E_i [\til P]$;
\item The system has \emph{a unique solution for $\wb$}  if whenever 
 $\til P$ and $\til Q$ are both solutions \hl{it holds that} $\til P \wb \til Q$. 
\end{itemize} 
 \end{definition}
\hl{Similarily, the \emph{(unique) solution of a system of equations for $\sim$}
(or for $\rapprox$) can be obtained by replacing all occurrences of $\wb$
in above definition with $\sim$ and $rapprox$, respectively.}

@@

\hl{Milner has totally three versions of ``unique solution of equations''
theorems, for $\sim$, $\wb$ and $\rapprox$, respectively, but only the
following two versions are explicitly mentioned} in~\cite{Mil89}:
\begin{theorem}[unique solution of equations for $\sim$]
\label{t:Mil89s1}
Let $E_i$ be weakly guarded with free variables $\subseteq \til X$,
and let ${\til P} \sim {\til E}\{\til P /\til X\}$,
  ${\til Q} \sim {\til E}\{\til Q /\til X\}$. Then ${\til P} \sim {\til Q}$.
\end{theorem}

\begin{theorem}[unique solution of equations for $\rapprox$]
\label{t:Mil89s3}
Let $E_i$ be guarded and sequential with free
variables $\subseteq \til X$, and let ${\til P} \rapprox {\til E}\{\til P /\til X\}$,
  ${\til Q} \rapprox {\til E}\{\til Q /\til X\}$. Then ${\til P} \rapprox {\til Q}$.
\end{theorem}

The version for $\wb$, due to the non-substitutivity of $\wb$, further requires
that all sums where equation variables appear are guarded sums:
\begin{theorem}[unique solution of equations for $\wb$]
\label{t:Mil89}
Let $E_i$ be guarded (and with guarded sums) and sequential with free
variables $\subseteq \til X$, and let ${\til P} \wb {\til E}\{\til P /\til X\}$,
  ${\til Q} \wb {\til E}\{\til Q /\til X\}$. Then ${\til P} \wb {\til Q}$.
\end{theorem}

The proof of Theorem~\ref{t:Mil89} exploits an invariance property on immediate transitions for
guarded and sequential expressions, and then extracts a bisimulation
(up to bisimilarity) out
of the solutions of the system.
To see the need of the sequentiality  condition, consider
 the equation (from \cite{Mil89}) $X = \res a (a. X | \outC a)$
where $X$ is guarded but not sequential. Any process that does not use
$a$ is a solution\hl{, e.g. $\nil$ and $b.\nil$.}

>>>>

\begin{definition}[equations] % Def 3.1
  \label{def:equation}
Assume that, for each $i$ of 
 a countable indexing set $I$, we have variables $X_i$, and expressions
$E_i$ containing variables up to $\til X$. Then 
$\{ X_i = E_i\}_{i\in I}$ is 
  a \emph{system of equations}. (There is one equation $E_i$ for each variable $X_i$.)
\end{definition}

@@

\begin{definition}[solutions and unique solutions]
  \label{def:solution}
Suppose $\{ X_i = E_i\}_{i\in I}$ is a system of equations: 
\begin{itemize}
\item
 $\til P$ is a \emph{solution of the system of equations (for $\wb$)} 
if for each $i$ it holds that $P_i \wb E_i [\til P]$;
\item The system has \emph{a unique solution for $\wb$}  if whenever 
 $\til P$ and $\til Q$ are both solutions \hl{it holds that} $\til P \wb \til Q$. 
\end{itemize} 
 \end{definition}
\hl{Similarily, the \emph{(unique) solution of a system of equations for $\sim$}
(or for $\rapprox$) can be obtained by replacing all occurrences of $\wb$
in above definition with $\sim$ and $rapprox$, respectively.}

@@

\hl{Milner has totally three versions of ``unique solution of equations''
theorems, for $\sim$, $\wb$ and $\rapprox$, respectively, but only the
following two versions are explicitly mentioned} in~\citep[p.~103, 158]{Mil89}:
\begin{theorem}[unique solution of equations for $\sim$]
\label{t:Mil89s1}
Let $E_i$ be weakly guarded with free variables up to $\til X$,
and let ${\til P} \sim {\til E}\{\til P /\til X\}$,
  ${\til Q} \sim {\til E}\{\til Q /\til X\}$. Then ${\til P} \sim {\til Q}$.
\end{theorem}

\begin{theorem}[unique solution of equations for $\rapprox$]
\label{t:Mil89s3}
Let $E_i$ be guarded and sequential with free
variables up to $\til X$, and let ${\til P} \rapprox {\til E}\{\til P /\til X\}$,
  ${\til Q} \rapprox {\til E}\{\til Q /\til X\}$. Then ${\til P} \rapprox {\til Q}$.
\end{theorem}

\hl{The version of Milner's unique-solution theorem for $\wb$ further requires
that all sums where equation variables appear are guarded
sums}\footnote{But if the CCS syntax were defined with only guarded
  sums, i.e., $\sum_{i\in I} \mu_i.P_i$ as
  in~\cite{sangiorgi2015equations}, this addition
  requirement disappears automatically, and we can even say $\wb$ is indeed a
  congruence.}:
\begin{theorem}[unique solution of equations for $\wb$]
\label{t:Mil89}
Let $E_i$ be guarded (with guarded sums only) and sequential with free
variables up to $\til X$, and let ${\til P} \wb {\til E}\{\til P /\til X\}$,
  ${\til Q} \wb {\til E}\{\til Q /\til X\}$. Then ${\til P} \wb {\til Q}$.
\end{theorem}

The proof of \hl{the last theorem above} exploits an invariance
property on immediate \hl{derivations}
of guarded and sequential expressions, and then extracts a bisimulation
(up to bisimilarity) out
of the solutions of the system.
To see the need of the sequentiality  condition, consider
 the equation (from \cite{Mil89}) $X = \res a (a. X | \outC a)$
where $X$ is guarded but not sequential. Any process that does not use
$a$ is a solution\hl{, e.g. $\nil$ and $b.\nil$.}

\hl{For more details of above three theorems,} c.f. Section~\ref{ss:part2}
for their formalised versions (the univariate case) and
Section~\ref{sec:multivariate} for the multivariate case
(Theorem~\ref{t:Mil89s1} only).

### 2019-09-09 ic-journal/unique.tex
Finished revising the first past

A \emph{system of contractions} is defined as a system of equations,
except that the contraction symbol $\mcontr$ is used in the place of
the equality symbol $=$. Thus a system of contractions is a set 
$\{  X_i \mcontr E_i\}_{i\in I}$
where $I$ is an  indexing set and expressions
$E_i$  may contain the  \behavC\  variables 
$\{  X_i\}_{i\in I}$.

\begin{definition}
\label{d:uniContra}
Given a system of contractions 
$\{  X_i \mcontr E_i\}_{i\in I}$, 
 we say that:
\begin{itemize}
\item $\til P$ is a \emph{solution (for $\mcontrBIS$) of the 
 system of contractions} if \hl{for each $i$ it holds that $P_i \mcontrBIS E_i [\til P]$};
\item \hl{The} system has \emph{a unique solution (for $\approx$)}
if $\til P \approx \til Q$ whenever $\til P$ and $\til Q$ are both solutions.
\end{itemize}
\end{definition}

The guardedness of contractions follows Def.~\ref{def:guardness} (for equations).
% \begin{definition}
% \label{d:guarded}
% A system of contractions $\{  X_i \mcontr E_i\}_{i\in I}$
%  is
% \emph{weakly guarded}
% if,  in each    $E_i$, each occurrence of
% a \behavC\ variable is underneath a prefix.

% The system use \emph{guarded sums} if 
% each $E_i$ only makes use of guarded sums.
% \end{definition}

@@

\begin{proof}{(sketch from \cite{sangiorgi2017equations})}
  Let $n$ be the length \hl{(the number of one-step transitions)} of
\hl{a} transition $\ct{\til P}\Arr\mu R$, and
let $\ctpp {\til P}$ and $\ctpp {\til Q}$  be the processes obtained
from  $\ct {\til P}$ and $\ct {\til Q}$ by unfolding the definitions
of the contractions $n$ times. Thus in $\qctpp$ each hole is
underneath at least $n$ prefixes, and cannot contribute to an action
in the first $n$ transitions; moreover all the contexts use guarded sums.

@@

\begin{theorem}[unique solution of contractions~\cite{sangiorgi2017equations}] % for $\wb$
\label{t:contraBisimulationU} % Thm 3.10
%A system of weakly-guarded contractions
%having only guarded sums, has a unique solution for $\wb$.
\hl{Let the expressions $E_i$ ($i \in I$) contain at most the
  variables $X_i$ ($i \in I$), and let each $X_j$ ($j\in I$) be weakly
  guarded in each $E_i$. For any two groups of CCS processes $P_i$ and
  $Q_i$ ($i \in I$), if $\til P \mcontrBIS \til E\{\til P/\til X\}$ and
    $\til Q \mcontrBIS \til E\{\til Q/\til X\}$ then $\til P \wb \til Q$.}
\end{theorem}

\begin{proof}
\hl{We want to prove $P_i \wb Q_i$ (for each $i \in I$) by considering} the following relation
\begin{equation*}
%\label{eq:R}
\R \DSdefi \{(R,S) \st R \wb \ct{\til P}, S \wb \ct{\til Q} \mbox{~for some context
$\qct$ (\hl{with only guarded sums})} \} \enspace.
\end{equation*}

@@

The unique solution theorem of Section~\ref{ss:SysContr} requires a
constrained syntax for sums, due to the congruence and precongruence
problems of bisimilarity and contraction with such operator. 
We show here that the constraints can be
removed by moving to the induced congruence and precongruence, the
latter called \emph{rooted contraction}:
\begin{definition}
\label{d:rcontra}
Two processes $P$ and $Q$ are in \emph{rooted contraction}, written as
 $P\rcontr Q$, if
\begin{enumerate}
\item $P \arr\mu P'$ implies that there is $Q'$ with $Q \arr \mu Q'$
 and $P'\mcontrBIS Q'$;
\item $Q \arr\mu Q'$   implies that there is $P'$ with $P \Arr \mu
 P'$ and $P' \wb Q'$\enspace.
\end{enumerate}
\end{definition}

%Above definition adapts the definition of rooted
%bisimilarity on top of that of the  contraction preorder
%$\mcontrBIS$.  %% Reviewer said this sentance is unclear. I too think so.

The precise formulation of  this definition was guided by the HOL theorem
prover and
the following two principles: (1) the definition should not be recursive,
along the lines of rooted bisimilarity
$\rapprox$ in Def.~\ref{d:rootedBisimilarity};
(2) the definition should  be built on top of existing \emph{contraction}
relation $\mcontrBIS$ (because of its completeness). 
A few other candidates were quickly tested and rejected, e.g.,
  because of  precongruence issue. The proof of the precongruence
 result below is along the lines of the analogous result
for rooted bisimilarity with respect to bisimilarity.

\begin{theorem}
\label{t:rcontrPrecongruence}
$\rcontr$ is a precongruence in CCS, and it is the
coarsest precongruence contained in $\contr$.
\end{theorem}

For a system of rooted contractions, the meaning of 
``solution for $\rcontr$'' and of \emph{a unique solution for $\rapprox$}
is the expected one --- just replace in Def.~\ref{d:uniContra}  the preorder 
$\contr$ with $\rcontr$, and the equivalence 
$\approx$ with $\rapprox$ (note that with $\rapprox$ one considers
solutions with respect to  $\rcontr$).
%
For this new relation, the analogous of Lemma~\ref{l:uptocon} and of
Theorem~\ref{t:contraBisimulationU} can now be stated without constraints on the sum
operator.
The schema of the proofs is almost identical, because all 
properties of $\rcontr$ needed in this proof is its precongruence, which is
indeed true on unrestricted contexts including direct sums:
\begin{lemma}
\label{l:ruptocon}
Suppose $\til P$ and $\til Q$ are solutions  for $\rcontr$ 
 of a system of weakly-guarded
contractions.
For any context $\qct$, 
if  $\ct{\til P}\Arr{\mu}  R$,
 then 
there is a  context $\qctp$
such that $R \mcontrBIS \ctp{\til P}$ and  $\ct{\til Q} \Arr{\mu} R'
 \wb \ctp{\til Q}$ \hl{for some $R'$}.
\end{lemma}
The proof of Lemma~\ref{l:ruptocon} is exactly the same as of
Lemma~\ref{l:uptocon}, \hl{because all we need from $\mcontrBIS$ (or
$\rcontr$) is their substitutivity properties (as precongruences)
w.r.t the corresponding definitions of contexts.}
Notice also that, the conclusion of
Lemma~\ref{l:ruptocon} still uses $\mcontrBIS$, as $\rcontr$ does
not hold.

\hl{The next lemma is actually Lemma 3.13} of~\citep[p.~102]{Mil89}, \hl{which
was designed to prove Milner's strong unique-solution theorem} (Prop.~3.14 of
\citep[p.~103]{Mil89}).
\begin{lemma}
  \label{lem:milner313}
  If the variables ${\til X}$ are weakly guarded in $E$, and $E\{\til
  P/\til X\} \arr{\alpha} P'$, then $P'$ takes the form $E'\{\til
  P/\til X\}$ (for some expression $E'$), and moreover, for any ${\til Q}$,
  $E\{\til Q/\til X\} \arr{\alpha} E'\{\til Q/\til X\}$.
\end{lemma}

\begin{theorem}[Unique solution of \hl{rooted} contractions]
\label{t:rcontraBisimulationU}
%A system of weakly-guarded contractions has a unique solution for $\rapprox$. 
% (thus also for $\wb$).
\hl{Let the expressions $E_i$ ($i \in I$) contain at most the
  variables $X_i$ ($i \in I$), and let each $X_j$ ($j\in I$) be weakly
  guarded in each $E_i$. For any two groups of CCS processes $P_i$ and
  $Q_i$ ($i \in I$), if $\til P \rcontr \til E\{\til P/\til X\}$ and
    $\til Q \rcontr \til E\{\til Q/\til X\}$ then $\til P \rapprox \til Q$.}
\end{theorem}

>>>>

A \emph{system of contractions} is defined as a system of equations,
except that the contraction symbol $\mcontr$ is used in the place of
$=$ in Def.~\ref{def:equation}. Thus a system of contractions is a set 
$\{  X_i \mcontr E_i\}_{i\in I}$ where $I$ is an indexing set and each
$E_i$  contains variables up to $\til X$.

\hl{Now we recall (with the proof) Sangiorgi's ``unique solution of
contractions'' theorem}~\cite{sangiorgi2017equations},
which greatly weakened the requirements of Milner's result
(Theorem~\ref{t:Mil89}).

@@

\begin{proof}{(sketch from~\cite{sangiorgi2017equations})}
  Let $n$ be the length \hl{(the number of one-step transitions)} of
\hl{a} transition $\ct{\til P}\Arr\mu R$, and
let $\ctpp {\til P}$ and $\ctpp {\til Q}$  be the processes obtained
from  $\ct {\til P}$ and $\ct {\til Q}$ by unfolding the definitions
of the contractions $n$ times. Thus in $\qctpp$ each hole is
underneath at least $n$ prefixes, and cannot contribute to an action
in the first $n$ transitions; moreover all the contexts use guarded sums.

@@

\begin{theorem}[unique solution of contractions~\cite{sangiorgi2017equations}]
\label{t:contraBisimulationU} % was: Thm 3.10
\hl{Let $E_i$ be weakly guarded (and with guarded sums only) with free
variables up to $\til X$, and let $\til P \mcontrBIS \til E\{\til P/\til X\}$,
    $\til Q \mcontrBIS \til E\{\til Q/\til X\}$. Then $\til P \wb \til Q$.}
\end{theorem}

\begin{proof}
\hl{We want to prove $P_i \wb Q_i$ (for each $i \in I$) by considering} the following relation
\begin{equation*}
\R \DSdefi \{(R,S) \st R \wb \ct{\til P}, S \wb \ct{\til Q} \mbox{~for some context
$\qct$ (\hl{with only guarded sums})} \} \enspace.
\end{equation*}

@@

Theorem~\ref{t:contraBisimulationU} brings a new proof technique for
bisimilarity, which is much less restrictive than Milner's
Theorem~\ref{t:Mil89} (but with the additional costs of checking $\mcontrBIS$
in addition to $\wb$).
However, comparing with Milner's Theorem~\ref{t:Mil89s1}, \hl{there is
still one more limit: the involved CCS contexts must only
use guarded sums. The is mainly due to the fact that $\wb$ is not a
 (real) congruence and also $\mcontrBIS$ is not a (real) precongruence.
Inspired by rooted bisimilarity, to completely eliminate the restriction
on guarded sums, we have extended the idea of contractions by moving
to \emph{rooted contractions}:}

\begin{definition}
\label{d:rcontra}
Two processes $P$ and $Q$ are in \emph{rooted contraction}, written as
 $P\rcontr Q$, if
\begin{enumerate}
\item $P \arr\mu P'$ implies that there is $Q'$ with $Q \arr \mu Q'$
 and $P'\mcontrBIS Q'$;
\item $Q \arr\mu Q'$   implies that there is $P'$ with $P \Arr \mu
 P'$ and $P' \wb Q'$\enspace.
\end{enumerate}
\end{definition}

\hl{The above definition was found with the help of interactive theorem proving.}
First of all, the following two principles were adopted when searching
for a possible definition: (1) the definition should not be \hl{coinductive},
along the lines of rooted bisimilarity $\rapprox$ (Def.~\ref{d:rootedBisimilarity});
(2) the definition should be built on top of existing \emph{contraction}
relation $\mcontrBIS$.
\hl{Then several candidates were quickly checked (using very similar
  proof steps): the key is to prove its transitivity, and whenever the
  proof attempts failed we tried to fix the definition. Eventually the
  above definition was found and shown to be a precongruence, i.e. it
  is preserved by all CCS operators except the Recursion, same as in
  the cases of $\sim$ and $\rapprox$.}
\hl{Furthermore, we needed to show that, the definition being found is the
best possible one, by showing that it is the coarsest precongruence
contains in $\mcontrBIS$, and the proof succeeded and is very similar
with the same result for $\rapprox$} (c.f.~Section~\ref{s:coarsest} for
more details).

\begin{theorem}
\label{t:rcontrPrecongruence}
$\rcontr$ is a precongruence in CCS, and it is the coarsest
precongruence contained in $\mcontrBIS$.
\end{theorem}

For this new relation, the analogous of Lemma~\ref{l:uptocon} and of
Theorem~\ref{t:contraBisimulationU} can now be stated without
constraints on the sum operator.
The schema of the proofs is almost identical, because all 
properties of $\rcontr$ needed in this proof is its precongruence, which is
indeed true on unrestricted contexts including direct sums:
\begin{lemma}
\label{l:ruptocon}
Suppose $\til P$ and $\til Q$ are solutions  for $\rcontr$ 
 of a system of weakly-guarded contractions.
For any context $\qct$, 
if  $\ct{\til P}\Arr{\mu}  R$,
 then there is a context $\qctp$
such that $R \mcontrBIS \ctp{\til P}$ and  $\ct{\til Q} \Arr{\mu} R'
 \wb \ctp{\til Q}$ \hl{for some $R'$}.
\end{lemma}
% The proof of Lemma~\ref{l:ruptocon} is exactly the same as of
% Lemma~\ref{l:uptocon}, \hl{because all we need from $\mcontrBIS$ (or
% $\rcontr$) is their substitutivity properties (as precongruences)
% w.r.t the corresponding definitions of contexts.}
% Notice also that, the conclusion of
% Lemma~\ref{l:ruptocon} still uses $\mcontrBIS$, as $\rcontr$ does
% not hold.

\hl{The next lemma is actually Lemma 3.13} of~\citep[p.~102]{Mil89}, \hl{which
was designed to prove Milner's ``unique solution of equations for
$\sim$''} (Theorem~\ref{t:Mil89s1}):
\begin{lemma}
  \label{lem:milner313}
  If the variables ${\til X}$ are weakly guarded in $E$, and $E\{\til
  P/\til X\} \arr{\alpha} P'$, then $P'$ takes the form $E'\{\til
  P/\til X\}$ (for some expression $E'$), and moreover, for any ${\til Q}$,
  $E\{\til Q/\til X\} \arr{\alpha} E'\{\til Q/\til X\}$.
\end{lemma}

\begin{theorem}[Unique solution of \hl{rooted} contractions]
\label{t:rcontraBisimulationU}
\hl{Let $E_i$ be weakly guarded with free variables up to $\til X$,
and let $\til P \rcontr \til E\{\til P/\til X\}$,
    $\til Q \rcontr \til E\{\til Q/\til X\}$. Then $\til P \rapprox \til Q$.}
\end{theorem}
